\renewcommand{\abstractname}{Abstract}
\begin{abstract}


* Hvorfor vi gjør det som vi gjør, hva er bakgrunnen

* Hva har vi gjort

* hvordan har vi gjort det

* hva fant vi ut

* hva resulterer det i hva konkluderer vi med?



Village Telco is an organization that aims to provide affordable communication in forms of data and voice services where no other companies can, or are willing to do so. Village Telco provides a “plug-and-play” solution with low cost voice and data service. While designed for the developing world, Village Telco’s solution can be applied anywhere where people wish to take control of their own telephone infrastructure.

This solution is delivered using an inexpensive fixed mesh WiFi delivery system called the Mesh Potato. The Mesh Potato unit is based on the open-source operating system, OpenWRT. Open Source telephony software combined with the latest wireless networking technology creates the potential for people to operate their own community phone systems. Mesh Potato networks have no dependence on existing telecom infrastructure, and can relatively easily be deployed anywhere in the world. It can either be deployed as a stand-alone solution or as an extension to existing technologies. Village Telco’s solution has been deployed in several countries around the world: from East-Timor and Nepal in Asia to several African and South America countries. The installed bases vary from 10 to several hundreds of Mesh Potatoes. 

We wanted to direct our studies toward the use of Mesh Potato in mobile situations. We started by looking into different scenarios consisting of everything from emergency situations and natural disasters to festivals and temporary refugee camps. After being in contact with different relief organization we got a clear impression that communication under and after a disaster situation is a difficult task. Keeith Williamson, a Village Telco volunteer, has created a Go-Box using the first generation of the Mesh Potato. We wanted to take his solution further and develop a mobile and stand-alone kit that could be used in all the different scenarios that we have taken into consideration. The prototype we developed differs from the one already made with the fact that we have used the second generation and that the box includes everything necessary, solar panel, battery, different cables and easy to use manuals for easy configuration. 

In the process of creating this kit we had to set up the network in order to fully test the network and the configurations. We found many of the descriptions to be little explanatory and difficult to use. Therefore a big part of our work has been to simplify these, and make them so easy that little or non technical people can use them. As well as come with suggestions and ideas to further simplifications. We have looked at simplifications as well as processes in order to make the roll-out as quick as possible. Hence we have named our solution the \gls{quick} box.

We believe that our work and theoretical research has contributed and richened Village Telco. And eventually save lives. 






\end{abstract}