\subsubsection{Script providing Internet access via PC}
\label{subsubsec:scriptvipc}

An alternative, and easier method in order to provide Internet access to the \gls{mp} via a PC with WiFi, is running a script. A script is a list of commands that can be executed with less user interactions to automate a process. The script is provided on a USB-stick that is included in the \gls{quick} box. 

A few easy steps must be conducted to run the script:
\begin{enumerate}
\item Connect the MP to the PC (make sure the MP is powered up) running Linux, with an Ethernet cable. The Ethernet cable must be plugged into the LAN-port on the MP.
\item Plug the provided USB-stick into the PC.  
\item Open the terminal on your PC by pressing "Ctrl+Alt+t". 
\item Execute the following commands to enter the right directory (where the script is located):
\noindent
\begin{lstlisting}[language=bash]
  $ sudo su
  $ cd /media/MP-USB
\end{lstlisting}
\item Execute the following command to run the script (Follow the descriptions provided in the script from now on. This information will both be given in the terminal window and as pop-up windows. Make sure you read these carefully):
\noindent
\begin{lstlisting}[language=bash]
  $ sh scriptviaPC.sh
\end{lstlisting}
\item After completing the steps given by the script, Internet access should be available in the mesh network. You can test this by connecting to the \gls{mp} from, for example, your smart phone or a PC. 
\begin{itemize}
\item The name of the network: \textbf{MP2_21}
\item The password is: \textbf{potato-potato}
\end{itemize}
Both the name and the password can be changed in the web interface. 
\end{enumerate}